---
title: "HW04"
author: "Oksana Ivanova"
date: "13 04 2019"
output: pdf_document
---

# Exploring dataset
```{r}
set.seed(42)
raw <- readRDS('weather.rds')
head(raw)
summary(raw)
str(raw)

```


# Problems of the dataset:

1) "X" collumn is just an index which is not needed for further analysis - have to be deleted.
2) Split [measure] column into multiple columns
3) It is better to assign our samples as [Year+month+day] from X number of columns.



```{r}
# Packages included
library(tidyr)
library(stringr)
# Get rid of [X] column
tidy <- raw[,-1]
head(tidy)
# Make dataframe from wide to long
tidy <- gather(tidy, day, value, X1:X31)
# Make dataframe from long to wide
tidy = spread(tidy, measure, value)
# Assign [Year+month+day] as samples names
tidy$day = str_sub(tidy$day, 2, 3)
tidy = unite(tidy, Date, year, month, day, sep = '_')
head(tidy)
```

4) Then it was noticed that some samples have all NAs data because of unexisting days (February 30, etc). 
   Also, the measurements in the end of the table are NAs as well.  


```{r}
tidy = na.omit(tidy)
```

5) Formatting of the numeric columns:

```{r}
tidy$PrecipitationIn = as.numeric(tidy$PrecipitationIn, na.string='T')
# Make columns with values as numbers numeric
tidy[,c(2,4:23)] = lapply(tidy[,c(2,4:23)], as.numeric)
# Check the dataframe after formatting
row.names(tidy) = 1:nrow(tidy)
summary(tidy)
```

6) Make [Events] column  as factor

```{r}

tidy$Events = as.factor(tidy$Events)
levels(tidy$Events)[1] = "Ordinary day"
head(tidy)
```

7) Plot the numeric columns data

```{r}
boxplot(tidy[, c(2, 4:23)])
```

8) Obvious outlier was detected - the [Humidity] was 1000% at one point of measurements that seems to be unreal and also [speedMPH] looks like outlier. I decided to change only [Humidity] data for this row for more realistic value = 100%.
 
```{r}
tidy[135,6] = 100
# Check the plot after additing
boxplot(tidy[, c(2, 4:23)])
```

# Finally, the dataset seems to be clear and tidy after processing and ready for further analysis.

Difference between "raw" and "tidy" dataframes.
 
```{r}
head(raw)
head(tidy)
```



